Stochastic Wind Power Forecasting

### 2020-03-10 LaTeX_code/paper/paper_1line.tex
Added Prof Kebaier theorem in apendex

\theoremstyle{definition}
\newtheorem{Def}{Definition}[section]
\newtheorem{Eg}{Example}[section]
\newtheorem{Rm}{Remark}[section]
\newtheorem{Ex}{Exercise}[section]
\newtheorem{Prop}[Def]{Proposition}
\newtheorem{Lem}[Def]{Lemma}
\newtheorem{Thm}[Def]{Theorem}
\newtheorem{Cor}[Def]{Corollary}
\newtheorem{Rem}[Def]{Remark}
\newtheorem{Ass}[Def]{Assumption}
\newtheorem{Met}[Def]{Method}
\newtheorem{Res}[Def]{Result}

@@

The  model becomes,
\begin{equation}
  \left\{
  \begin{array}{@{}rl@{}}
    dV_t \!\!\!&=  - \theta_t V_t \  dt + \sqrt{2 \theta_t \alpha (V_t +p_t ) (1-V_t-p_t)} \  dW_t   \\ %\quad t > 0
V_0 \!\!\!&=  v_0 \,. \label{VtSDE}
% \theta_t &= \max \left( \theta_0 \ , \ \frac{|\dot{p}_t|}{\min (p_t, 1-p_t)}  \right )\\
\end{array}\right. 
\end{equation}
%---END SECTION 2---

@@

 By assuming ergodicity {\color{red} not ergodicity, instead simply for the quadratic variation formula.}, we can obtain a first estimate on the product of the parameters as follows,

@@

\begin{Thm}
Assume that assumptions of Theorem \ref{thm:exun} hold with $x_0\in]0,1[$.
Let $\tau_0:=\inf \{t\in[0,T],\; X_t=0\}$ and  $\tau_1:=\inf \{t\in[0,T],\; X_t=1\}$ with the convention that $\inf\emptyset=+\infty$. Assume in addition that
\begin{equation}\label{Assumption:2}
0<\alpha < \frac 12\quad \mbox{ and }\quad \forall  t\in[0,T],\;\; \alpha\theta_t\le \dot p_t +\theta_tp_t \le (1-\alpha)\theta_t \tag{B}. 
\end{equation}
 Then, $\tau_0=\tau_1=+\infty$ a.s.
\end{Thm}
\begin{proof}
For $t\in[0,\tau_0[$, we have 
$$
\frac{dX_t}{X_t}=\left( \frac{\dot p_t +\theta_t p_t}{X_t} - \theta_t\right)\dif t  +\sqrt{\frac{2\alpha \theta_t(1-X_t)}{X_t}} \dif W_t 
$$  
so that
$$
X_t=x_0\exp\Big(\int_0^t \frac{\dot p_s +\theta_s(p_s-\alpha)}{X_s}ds-(1-\alpha)\int_0^t\theta_sds + M_t\Big),
$$
where $M_t=\int_0^t\sqrt{\frac{2\alpha \theta_s(1-X_s)}{X_s}} dW_s$ is a continuous martingale. Then as for all $t\in[0,T]$, we have $\dot p_t +\theta_t(p_t-\alpha)\ge0$, we deduce that
$$
X_t\ge x_0\exp\Big(-(1-\alpha)\int_0^t\theta_sds + M_t\Big).
$$
By way of contradiction let us assume that  $\{\tau_0<\infty\}$, then letting $t\to \tau_0$ we deduce that $\lim_{t\to \infty} \mathbf 1_{\{\tau_0<\infty\}}M_{t\wedge \tau_0}=\mathbf -1_{\{\tau_0<\infty\}}\infty$ a.s. This leads to a contradiction since we know that continuous martingales likewise the Brownian motion cannot converge  almost surely to $+\infty$ or $-\infty$. It follows that $\tau_0=\infty$ almost surely. Next, recalling that  the process $(Y_t)_{t\geq 0}$  given by $Y_t=1-X_t$ is solution to 
$$
dY_t= \big( \theta_t(1-p_t) -\dot p_t - \theta_tY_t  \big)dt  -\sqrt{2\alpha \theta_tY_t(1-Y_t)} dW_t 
$$
we deduce using similar arguments as above 
$\tau_1=\infty$ a.s. provided that $\theta_t(1-p_t) -\dot p_t -\alpha \theta_t\ge 0$.
 \end{proof}
\textcolor{red}{When using Lamperti, the form }

>>>>

\theoremstyle{definition}
\newtheorem{Def}{Definition}[section]
\newtheorem{Eg}{Example}[section]
\newtheorem{Rm}{Remark}[section]
\newtheorem{Ex}{Exercise}[section]
\newtheorem{Prop}[Def]{Proposition}
\newtheorem{Lem}[Def]{Lemma}
\newtheorem{Thm}[Def]{Theorem}
\newtheorem{Cor}[Def]{Corollary}
\newtheorem{Rem}[Def]{Remark}
\newtheorem{Ass}[Def]{Assumption}
\newtheorem{Met}[Def]{Method}
\newtheorem{Res}[Def]{Result}
\newcommand{\red}{\textcolor{red}}

@@

The  model becomes,
\begin{equation}
\begin{cases}
    \dif V_t &=  - \theta_t V_t \ \dif t + \sqrt{2 \theta_t \alpha (V_t +p_t ) (1-V_t-p_t)} \  \dif W_t \\
V_0 \!\!\!&=  v_0 \,.\\ 
%\theta_t &= \max \left( \theta_0 \ , \ \frac{|\dot{p}_t|}{\min (p_t, 1-p_t)}  \right )
\end{cases}
\label{VtSDE}
\end{equation}
%---END SECTION 2---

@@

 By assuming ergodicity {\color{red} not ergodicity, instead simply for the \textbf{quadratic variation formula}.}, we can obtain a first estimate on the product of the parameters as follows,

@@

\begin{Thm}
Assume that assumptions of Theorem \ref{thm:exun} hold with $x_0\in]0,1[$.
Let $\tau_0:=\inf \{t\in[0,T],\; X_t=0\}$ and  $\tau_1:=\inf \{t\in[0,T],\; X_t=1\}$ with the convention that $\inf\emptyset=+\infty$. Assume in addition that
\begin{equation}\label{Assumption:2}
0<\alpha < \frac 12\quad \mbox{ and }\quad \forall  t\in[0,T],\;\; \alpha\theta_t\le \dot p_t +\theta_tp_t \le (1-\alpha)\theta_t \tag{B}. 
\end{equation}
 Then, $\tau_0=\tau_1=+\infty$ a.s.
\end{Thm}
\begin{proof}
For $t\in[0,\tau_0[$, we have 
$$
\frac{dX_t}{X_t}=\left( \frac{\dot p_t +\theta_t p_t}{X_t} - \theta_t\right)\dif t  +\sqrt{\frac{2\alpha \theta_t(1-X_t)}{X_t}} \dif W_t 
$$  
so that
$$
X_t=x_0\exp\Big(\int_0^t \frac{\dot p_s +\theta_s(p_s-\alpha)}{X_s}ds-(1-\alpha)\int_0^t\theta_sds + M_t\Big),
$$
where $M_t=\int_0^t\sqrt{\frac{2\alpha \theta_s(1-X_s)}{X_s}} dW_s$ is a continuous martingale. Then as for all $t\in[0,T]$, we have $\dot p_t +\theta_t(p_t-\alpha)\ge0$, we deduce that
$$
X_t\ge x_0\exp\Big(-(1-\alpha)\int_0^t\theta_sds + M_t\Big).
$$
By way of contradiction let us assume that  $\{\tau_0<\infty\}$, then letting $t\to \tau_0$ we deduce that $\lim_{t\to \infty} \mathbf 1_{\{\tau_0<\infty\}}M_{t\wedge \tau_0}=\mathbf -1_{\{\tau_0<\infty\}}\infty$ a.s. This leads to a contradiction since we know that continuous martingales likewise the Brownian motion cannot converge  almost surely to $+\infty$ or $-\infty$. It follows that $\tau_0=\infty$ almost surely. Next, recalling that  the process $(Y_t)_{t\geq 0}$  given by $Y_t=1-X_t$ is solution to 
$$
dY_t= \big( \theta_t(1-p_t) -\dot p_t - \theta_tY_t  \big)dt  -\sqrt{2\alpha \theta_tY_t(1-Y_t)} dW_t 
$$
we deduce using similar arguments as above 
$\tau_1=\infty$ a.s. provided that $\theta_t(1-p_t) -\dot p_t -\alpha \theta_t\ge 0$.
\end{proof}

\section{New model}
In what follows we adapt the proof of the above theorem for the new model 
\begin{align}\label{eq:model2}
dX_t&= \big(\dot p_t - \theta_t(X_t-p_t)  \big)dt  +\sqrt{2\alpha \red{\theta_0}X_t(1-X_t)} dW_t\quad t\in[0,T]\\
X_0&=x_0\in[0,1],\notag
\end{align}

\begin{Thm}\label{thm:mod2}
Assume that assumptions of Theorem \ref{thm:exun} hold with $x_0\in]0,1[$.
Let $\tau_0:=\inf \{t\in[0,T],\; X_t=0\}$ and  $\tau_1:=\inf \{t\in[0,T],\; X_t=1\}$ with the convention that $\inf\emptyset=+\infty$. Assume in addition that for all $t\in[0,T]$,  $p_t\in]0,1[$ and that
\red{
\begin{equation}\label{Assumption:3}
\theta_t\geq \max\left(\frac{\alpha\theta_0+\dot p_t}{1-p_t},\frac{\alpha\theta_0-\dot p_t}{p_t}\right)\tag{C}. 
\end{equation}
}
 Then, $\tau_0=\tau_1=+\infty$ a.s.
\end{Thm}

\begin{proof}
For $t\in[0,\tau_0[$, we have 
$$
\frac{dX_t}{X_t}= \left(\frac{\dot p_t +\theta_t p_t}{X_t} - \theta_t\right)dt  +\sqrt{\frac{2\alpha \red{\theta_0}(1-X_t)}{X_t}} dW_t 
$$  
so that
$$
X_t=x_0\exp\Big(\int_0^t \frac{\dot p_s +\theta_sp_s- \red{\theta_0}\alpha)}{X_s}ds+\alpha\red{\theta_0}t-  \int_0^t\theta_sds + M_t\Big),
$$
where $M_t=\int_0^t\sqrt{\frac{2\alpha \theta_s(1-X_s)}{X_s}} dW_s$ is a continuous martingale. Then as for all $t\in[0,T]$, we have $\dot p_t +\theta_tp_t- \red{\theta_0}\alpha\ge0$, we deduce that
$$
X_t\ge x_0\exp\Big(\alpha\red{\theta_0}t-  \int_0^t\theta_sds + M_t\Big).
$$
By way of contradiction let us assume that  $\{\tau_0<\infty\}$, then letting $t\to \tau_0$ we deduce that $\lim_{t\to \infty} \mathbf 1_{\{\tau_0<\infty\}}M_{t\wedge \tau_0}=\mathbf -1_{\{\tau_0<\infty\}}\infty$ a.s. This leads to a contradiction since we know that continuous martingales likewise the Brownian motion cannot converge  almost surely to $+\infty$ or $-\infty$. It follows that $\tau_0=\infty$ almost surely. Next, recalling that  the process $(Y_t)_{t\geq 0}$  given by $Y_t=1-X_t$ is solution to 
$$
dY_t= \big( \theta_t(1-p_t) -\dot p_t - \theta_tY_t  \big)dt  -\sqrt{2\alpha \red{\theta_0}Y_t(1-Y_t)} dW_t 
$$
we deduce using similar arguments as above 
$\tau_1=\infty$ a.s. provided that $\theta_t(1-p_t) -\dot p_t -\alpha \red{\theta_0}\ge 0$.
 \end{proof}
 \section{Answers to Renzo's questions} 
{\color{blue}So, Theorem \ref{thm:mod2} answers the second question in slide 11.} The new condition \eqref{Assumption:3} that guarantees  for $(X_t)_{t\ge0}$ to stay in $]0,1[$ depends now on $\alpha$. I would keep condition \eqref{Assumption:3} as we have a proof for it.  Moreover, if condition \eqref{Assumption:3} is satisfied, then  the orange condition in slide 8 is also satisfied 
since we have 
$$
\max\left(\frac{\alpha\theta_0+\dot p_t}{1-p_t},\frac{\alpha\theta_0-\dot p_t}{p_t}\right)\ge \max\left(\frac{\frac{\alpha\theta_0}{2}+\dot p_t}{1-p_t},\frac{\frac{\alpha\theta_0}{2}-\dot p_t}{p_t}\right)
$$
which ensures that $Z$ don't hit the boundaries too according to Renzo's intuitive arguments.\\ 
{\color{blue}Concerning, the first question on slide 11:} we can't apply Itô's formula, unless we are sure that the process don't hit the boundaries, that's why we have to assume  \eqref{Assumption:3} before using Itô's formula.  In addition, as the diffusion coefficient of $X$  given by  $x\mapsto \sqrt{\frac{2\alpha \red{\theta_0}(1-x)}{x}}$  is strictly positive for all $x$ in the state space of $X$, namely $]0,1[$ under assumption \eqref{Assumption:3},  this is enough to ensure that the transformation between $Z$ and $X$ is bijective so that we deduce the existence and uniqueness of $Z$ from those of $X$. Otherwise the Lamperti transform is not valid if the state space of $X$ is $[0,1]$.
